Renewal and stability in populations structured by remaining years of life

### 2014-10-20 ThanoRepro/ThanoRepro.tex
tinkering with tikz, won't compile on refugee server due to old ubuntu version...

\documentclass{article}
\usepackage{appendix}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{placeins}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tikz}
%\usepackage[active,tightpage]{preview}
\usepackage{natbib}
\bibpunct{(}{)}{,}{a}{}{;} 
\usepackage{url}
\usepackage{nth}
% for the d in integrals
\newcommand{\dd}{\; \mathrm{d}}
\newcommand{\ec}{\quad\quad\text{,}}
\newcommand{\ep}{\quad\quad\text{.}}

\usepackage[top=2in, bottom=1.5in, left=1in, right=1in]{geometry}
\usepackage{setspace}
% matrix numbering
\usepackage{trivfloat}
\trivfloat{matrix}
\floatstyle{plaintop}
    \restylefloat{matrix}
\AtBeginDocument{\numberwithin{matrix}{section}}
% end preamble
%-------------------------------------------------------

\begin{document}

@@

\begin{figure}[ht!]
	\caption{2010 US population structure}
	\begin{center}
	\begin{subfigure}{.45\textwidth}
		\caption{Chronological age-structure (years lived) decomposed by
		thanatological age (years left).}
		\label{fig:USdecomp}
		\includegraphics[scale=.6]{Figures/US2010Age.pdf}
	\end{subfigure}
	~
	\begin{subfigure}{.45\textwidth}
		\caption{Thanatological age-structure (years left) decomposed by chronological
		age (years lived).}
		\label{fig:USrecomp}
		\includegraphics[scale=.6]{Figures/US2010Thano.pdf}
	\end{subfigure}
	\\
	\small{*Population and mortality data from the HMD.}
	\end{center}
\end{figure}

@@

We are interested to use thanatological fertility rates, $f^\star(y)$ in
models of population renewal,\footnote{This is probably premature, but helps to
understand the nature and provenance of model components.} rather than to
explicitly study the nature of these rates. Nonetheless it is best not to throw such rates into a model blindly, so we offer a schematic overview of some of their characteristics.
Figure~\ref{fig:FySpaghetti} represents the full variety in
female thanatological period fertility rates that can be found for all years
of data that overlap in the HFD and HMD.\footnote{There are as of this writing
1834 population-years of overlap between the HMD and HFD, including a wide
variety of fertility and mortality combinations.} Figure~\ref{fig:FxSpaghetti} gives ASFR for the same
populations and years as a more familiar reference, but note that both scales
are different. With no indication of particular populations or time series, one
already concludes that period thanatological fertility rates have a
characteristic shape and are not random, erratic or informationless: There is a
pattern to fertility rates by remaining years of life, and it varies over time
and between populations within some range of normality. Thanatological fertility
rates have a wider distribution than do chronological rates. Note that this
spaghetti plot includes both fertility booms and busts, as well as some
mortality crises (1918, WWII). Some patterns to note, but which we do not
separate graphically here, are that the left tail (a gauge of how orphan-prone a
population is) has tended to fall over time. All populations have shown a
rightward shift in both the mean and modal mother's thanatological age at birth
(TAB) over time. Several
populations now show modal TABs of over 60 years. Thanatological total fertility rates (TTFR)
track standard period total fertility rates (PTFR) rather closely, but tend to
be somewhat higher (not shown).

\begin{figure}[h!]
	\caption{Chronological and thantological fertility rates, all 1600
	country-year combinations present in both the HFD and HMD.* Note different x
	and y scales.}
	\label{fig:Fxcompare}
	\begin{center}
	\makebox[\textwidth]{
	\begin{subfigure}{.45\textwidth}
		\caption{Chronological period fertility rates.}
		\label{fig:FxSpaghetti}
		\includegraphics[scale=.45]{Figures/FxSpaghettiDraft.png}
	\end{subfigure}
	~
	\begin{subfigure}{.45\textwidth}
		\caption{Thanatological period fertility rates.}
		\label{fig:FySpaghetti}
		\includegraphics[scale=.45]{Figures/FySpaghettiDraft.png}
	\end{subfigure}
	}
	\\
	\end{center}
	\begin{tiny}
	*AUT, $1951-2010$; BGR, $1947-2009$; BLR, $1964-2009$; CAN, $1921-2007$; 
	CHE, $1932-2011$; CZE, $1950-2011$; DEUTE, $1956-2010$; DEUTNP, $1990-2010$; 
	DEUTW, $1956-2010$; ESP, $1922-2006$; EST, $1959-2010$; FIN, $1939-2009$; 
	FRATNP, $1946-2010$; GBR\_NIR, $1974-2009$; GBR\_NP, $1974-2009$; GBR\_SCO,
	$1945-2009$; GBRTENW, $1938-2009$; HUN, $1950-2009$; IRL, $1955-2009$; JPN, $1947-2009$; 
	LTU, $1959-2010$; NLD, $1950-2009$; NOR, $1967-2009$; PRT, $1940-2009$; 
	RUS, $1959-2010$; SVK, $1950-2009$; SVN, $1983-2009$; SWE, $1891-2010$; 
	TWN, $1976-2010$; UKR, $1959-2006$; USA, $1933-2010$
	\end{tiny}
\end{figure}

@@

\begin{figure}
\centering
\caption{The thanatological lifecycle graph}
\label{fig:graph}
\begin{tikzpicture}[->,shorten >=1pt,auto,node distance=4.3cm,
  thick,main node/.style={circle,fill=blue!20,draw,font=\sffamily\Large\bfseries}]
  \definecolor{gray}{rgb}{0.8,0.8,0.8}
  \definecolor{lightgreen}{rgb}{0,0.8,0.8}
  \node[main node] (0) {0};
  \node[main node] (5) [above right of=0] {5};
  \node[main node] (4) [right of=5] {4};
  \node[main node] (3) [below right of=4] {3};
  \node[main node] (2) [below left of=3] {2};
  \node[main node] (1) [left of=2] {1};
  \node[main node] (X) [left of = 5] {X};

@@

    \path[every node/.style={font=\sffamily\small}, ultra thick, color =
    lightgreen ] (5) edge node {} (4)
    (4) edge node {} (3)
    (3) edge node {} (2)
    (2) edge node {} (1)
    (1) edge node {} (0)
    (0) edge node {} (X);

\end{tikzpicture}
\end{figure}

>>>>

\documentclass{article}
\usepackage{appendix}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{placeins}
\usepackage{graphicx}
\usepackage{subcaption}
%\usepackage{pgf}
\usepackage{tikz}
%\usetikzlibrary{arrows}
%\usepackage[active,tightpage]{preview}
%\PreviewEnvironment{tikzpicture}
%\setlength\PreviewBorder{5pt}
\usepackage{natbib}
\bibpunct{(}{)}{,}{a}{}{;} 
\usepackage{url}
\usepackage{nth}
% for the d in integrals
\newcommand{\dd}{\; \mathrm{d}}
\newcommand{\ec}{\quad\quad\text{,}}
\newcommand{\ep}{\quad\quad\text{.}}

\usepackage[top=2in, bottom=1.5in, left=1in, right=1in]{geometry}
\usepackage{setspace}
% matrix numbering
\usepackage{trivfloat}
\trivfloat{matrix}
\floatstyle{plaintop}
    \restylefloat{matrix}
\AtBeginDocument{\numberwithin{matrix}{section}}
% end preamble
%-------------------------------------------------------
\begin{document}

@@

% both figures created in R/Pyramids.R
\begin{figure}[ht!]
	\caption{2010 US population structure}
	\begin{center}
	\begin{subfigure}{.45\textwidth}
		\caption{Chronological age-structure (years lived) decomposed by
		thanatological age (years left).}
		\label{fig:USdecomp}
		\includegraphics[scale=.6]{Figures/US2010Age.pdf}
	\end{subfigure}
	~
	\begin{subfigure}{.45\textwidth}
		\caption{Thanatological age-structure (years left) decomposed by chronological
		age (years lived).}
		\label{fig:USrecomp}
		\includegraphics[scale=.6]{Figures/US2010Thano.pdf}
	\end{subfigure}
	\\
	\small{*Population and mortality data from the HMD.}
	\end{center}
\end{figure}

@@

We are interested to use thanatological fertility rates, $f^\star(y)$ in
models of population renewal,\footnote{This is probably premature, but helps to
understand the nature and provenance of model components.} rather than to
explicitly study the nature of these rates. Nonetheless it is best not to throw such rates into a model blindly, so we offer a schematic overview of some of their characteristics.
Figure~\ref{fig:FySpaghetti} represents the full variety in
female thanatological period fertility rates that can be found for all years
of data that overlap in the HFD and HMD.\footnote{There are as of this writing
1834 population-years of overlap between the HMD and HFD, including a wide
variety of fertility and mortality combinations.} Figure~\ref{fig:FxSpaghetti} gives ASFR for the same
populations and years as a more familiar reference, but note that both scales
are different. With no indication of particular populations or time series, one
already concludes that period thanatological fertility rates have a
characteristic shape and are not random, erratic or informationless: There is a
pattern to fertility rates by remaining years of life, and it varies over time
and between populations within some range of normality. Thanatological fertility
rates have a wider distribution than do chronological rates. Note that this
spaghetti plot includes both fertility booms and busts, as well as some
mortality crises (1918, WWII). Some patterns to note, but which we do not
separate graphically here, are that the left tail (a gauge of how orphan-prone a
population is) has tended to fall over time. All populations have shown a
rightward shift in both the mean and modal mother's thanatological age at birth
(TAB) over time. Several
populations now show modal TABs of over 60 years. Thanatological total fertility rates (TTFR)
track standard period total ferti
lity rates (PTFR) rather closely, but tend to
be somewhat higher (not shown).
% both figures created in R/HFDspaghetti.R
\begin{figure}[h!]
	\caption{Chronological and thantological fertility rates, all 1600
	country-year combinations present in both the HFD and HMD.* Note different x
	and y scales.}
	\label{fig:Fxcompare}
	\begin{center}
	\makebox[\textwidth]{
	\begin{subfigure}{.45\textwidth}
		\caption{Chronological period fertility rates.}
		\label{fig:FxSpaghetti}
		\includegraphics[scale=.45]{Figures/FxSpaghettiDraft.png}
	\end{subfigure}
	~
	\begin{subfigure}{.45\textwidth}
		\caption{Thanatological period fertility rates.}
		\label{fig:FySpaghetti}
		\includegraphics[scale=.45]{Figures/FySpaghettiDraft.png}
	\end{subfigure}
	}
	\\
	\end{center}
	\begin{tiny}
	*AUT, $1951-2010$; BGR, $1947-2009$; BLR, $1964-2009$; CAN, $1921-2007$; 
	CHE, $1932-2011$; CZE, $1950-2011$; DEUTE, $1956-2010$; DEUTNP, $1990-2010$; 
	DEUTW, $1956-2010$; ESP, $1922-2006$; EST, $1959-2010$; FIN, $1939-2009$; 
	FRATNP, $1946-2010$; GBR\_NIR, $1974-2009$; GBR\_NP, $1974-2009$; GBR\_SCO,
	$1945-2009$; GBRTENW, $1938-2009$; HUN, $1950-2009$; IRL, $1955-2009$; JPN, $1947-2009$; 
	LTU, $1959-2010$; NLD, $1950-2009$; NOR, $1967-2009$; PRT, $1940-2009$; 
	RUS, $1959-2010$; SVK, $1950-2009$; SVN, $1983-2009$; SWE, $1891-2010$; 
	TWN, $1976-2010$; UKR, $1959-2006$; USA, $1933-2010$
	\end{tiny}
\end{figure}

@@

\begin{figure}
\centering
\caption{The thanatological lifecycle graph}
\label{fig:graph}
\begin{tikzpicture}[->,shorten>=1pt,auto,node distance=4.3cm,thick,main
node/.style={circle,fill=blue!20,draw,font=\sffamily\Large\bfseries}]
  \definecolor{gray}{rgb}{0.8,0.8,0.8}
  \definecolor{lightgreen}{rgb}{0,0.8,0.8}

  \node[main node] (0) {0};
  \node[main node] (5) [above right of=0] {5};
  \node[main node] (4) [right of=5] {4};
  \node[main node] (3) [below right of=4] {3};
  \node[main node] (2) [below left of=3] {2};
  \node[main node] (1) [left of=2] {1};
  \node[main node] (X) [left of = 5] {X};

@@

    \path[every node/.style={font=\sffamily\small}, ultra thick, color=lightgreen ] 
    (5) edge node {} (4)
    (4) edge node {} (3)
    (3) edge node {} (2)
    (2) edge node {} (1)
    (1) edge node {} (0)
    (0) edge node {} (X);    
\end{tikzpicture}
\end{figure}
